# Ascertainment

## Example

Burton et al (2000) provide the following example:

They simulated a population containing 4,800 sibships of size 3, with a total of 14,400 children, and 864 affected children. In this population, the prevalence of the disease is 864/14400, which equals 0.06:

```{r}
864/14400
```

In their simulation, they then sampled every family that contained one or more affected child, and ended up with this set of 813 families:

N Aff Children | N Sibships 
------------- | ------------- 
1 | 763 
2 | 49 
3 | 1

**Question 1**: What is the prevalence in this sample of 813 families if we compute it without employing any ascertainment correction?

```{webr-r}
# Edit/add R code here
         
```

::: {.callout-tip collapse="true"}
## Answer 1

The prevalence without any ascertainment correction is 864/(3\*813) = 0.3542.

```{r}
# Total number of affecteds
763 + 2*49 + 3*1
# Total number of children
3*813
# (Number of affecteds)/(Number of children)
864/(3*813)
```
:::

**Question 2**: What is the prevalence estimate after correcting for ascertainment?

```{webr-r}
# Edit/add R code here
        
```

::: {.callout-tip collapse="true"}
## Answer 2

Here we use the Li and Mantel (1968) formula

p = (R - J)/(T - J)

where

-   R = the total number of affected siblings
-   T = the total number of siblings
-   J = the number of siblings with only one proband

In this case,

-   R = 864
-   T = 3 \* 813 = 2439
-   J = 763

So p = (R - J)/(T - J) = (864 - 763)/(2439 - 763) = 0.0603

```{r}
R = 864
T = 3 * 813 
J = 763
(R - J)/(T - J)
```
:::

## References

1.  Burton PR, Palmer LJ, Jacobs K, Keen KJ, Olson JM, Elston RC. Ascertainment adjustment: where does it take us? Am J Hum Genet. 2000 Dec;67(6):1505–1514. PMID: 11078478 PMCID: PMC1287927 DOI: https://doi.org/10.1086/316899

2.  Li CC, Mantel N. A simple method of estimating the segregation ratio under complete ascertainment. Am J Hum Genet. 1968 Jan;20(1):61–81. PMID: 5635671 PMCID: PMC1706251

3.  Davie AM. The “singles” method for segregation analysis under incomplete ascertainment. Ann Hum Genet. 1979 May;42(4):507–512. PMID: 475337 DOI: https://doi.org/10.1111/j.1469-1809.1979.tb00683.x


